---
title: "DATA 621 Assignment 1"
author: "Philip Tanofsky"
date: "9/13/2020"
output: html_document
---

## Assignment 1

The goal of this assignment is to create a linear regression model of baseball team statistics to predict the number of wins for a given team.


Following example: http://www.sthda.com/english/articles/40-regression-analysis/168-multiple-linear-regression-in-r/#:~:text=Multiple%20linear%20regression%20is%20an,distinct%20predictor%20variables%20(x).&text=The%20%E2%80%9Cb%E2%80%9D%20values%20are%20called,weights%20(or%20beta%20coefficients).

Following approach here, too: https://machinelearningmastery.com/machine-learning-in-r-step-by-step/

```{r libraries}
library(tidyverse)
library(ggplot2)
library(GGally)
library(caret)
library(ggExtra)
```


```{r}
# Read in CSV file of training dat
mb_train_init <- read.csv("moneyball-training-data.csv")
```

```{r}
# Dimensions of the dataset
dim(mb_train_init)
```

- 2276 instances
- 17 variables (1 index, 1 dependent variable, 15 independent variables)

```{r}
# list types of each attribute
sapply(mb_train_init, class)
```

- Data type for every variable is integer. Makes sense given the statistical nature of the data. Statistics are expected to be whole number integers.

```{r}
# first five rows
head(mb_train_init)
```

Outputting first 5 instances for simple visual inspection. Initial observation, TEAM_BATTING_HBP is NA (not available) for each instance. Will need further investigation.

```{r}
# summary
summary(mb_train_init)
```

Display summary of all the variables (statistics)

#### Variables containing at least one NA value

- TEAM_BATTING_SO
- TEAM_BASERUN_SB
- TEAM_BASERUN_CS
- TEAM_BATTING_HBP (majority of instances contain NA)
- TEAM_PITCHING_SO
- TEAM_FIELDING_DP

#### Interesting observations

- Minimum number of wins is 0: Given a 162 game schedule, I don't believe 0 wins is a realistic value
- Maximum number of wins is 146: Given a 162 game schedule, this is an unrealistic number. Most wins by a MLB team is 116. 1906 Chicago Cubs and 2001 Seattle Mariners
- Maximum number of TEAM_PITCHING_H is 30132: This indicates a team gave up over 180 hits per game. That's completely urealistic.
- Maximum number of TEAM_PITCHING_SO is 19278: This indicates a team achieved 119 striketouts per game. Literally impossible.


```{r}
# More Visualize
ggpairs(mb_train_init[,3:7], color="gray20")
```

Not much value in above visualization

```{r}
p1 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_H, y=TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method=lm) + 
  labs(x = "Batting Hits", y = "Wins", title="Wins by Total Hits")
```

```{r}
# Scatterplot with density plot
ggMarginal(p1, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p1, type = "boxplot")
```

```{r}
p2 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_2B, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting Doubles", y = "Wins", title="Wins by Doubles")
```

```{r}
# Scatterplot with density plot
ggMarginal(p2, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p2, type = "boxplot")
```

```{r}
p3 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_3B, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting Triples", y = "Wins", title="Wins by Triples")
```

```{r}
# Scatterplot with density plot
ggMarginal(p3, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p3, type = "boxplot")
```

```{r}
p4 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_HR, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting Home Runs", y = "Wins", title="Wins by Home Runs")
```

```{r}
# Scatterplot with density plot
ggMarginal(p4, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p4, type = "boxplot")
```

```{r}
p5 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_BB, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting Walks", y = "Wins", title="Wins by Walks")
```

```{r}
# Scatterplot with density plot
ggMarginal(p5, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p5, type = "boxplot")
```

```{r}
p6 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_SO, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting Strikeouts", y = "Wins", title="Wins by Strikeouts")
```

```{r}
# Scatterplot with density plot
ggMarginal(p6, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p6, type = "boxplot")
```

```{r}
p7 <- ggplot(mb_train_init, aes(x=TEAM_BATTING_HBP, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Batting HBP", y = "Wins", title="Wins by HBP")
```

```{r}
# Scatterplot with density plot
ggMarginal(p7, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p7, type = "boxplot")
```

```{r}
p8 <- ggplot(mb_train_init, aes(x=TEAM_BASERUN_SB, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Base running Stolen Bases", y = "Wins", title="Wins by Stolen Bases")
```

```{r}
# Scatterplot with density plot
ggMarginal(p8, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p8, type = "boxplot")
```

```{r}
p9 <- ggplot(mb_train_init, aes(x=TEAM_BASERUN_CS, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Base running Caught Stealing", y = "Wins", title="Wins by Caught Stealing")
```

```{r}
# Scatterplot with density plot
ggMarginal(p9, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p9, type = "boxplot")
```

```{r}
p10 <- ggplot(mb_train_init, aes(x=TEAM_PITCHING_H, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Pitching Hits Allowed", y = "Wins", title="Wins by Hits Allowed")
```

```{r}
# Scatterplot with density plot
ggMarginal(p10, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p10, type = "boxplot")
```

```{r}
p11 <- ggplot(mb_train_init, aes(x=TEAM_PITCHING_HR, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Pitching Home Runs Allowed", y = "Wins", title="Wins by Home Runs Allowed")
```

```{r}
# Scatterplot with density plot
ggMarginal(p11, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p11, type = "boxplot")
```

```{r}
p12 <- ggplot(mb_train_init, aes(x=TEAM_PITCHING_BB, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Pitching Walks Allowed", y = "Wins", title="Wins by Walks Allowed")
```

```{r}
# Scatterplot with density plot
ggMarginal(p12, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p12, type = "boxplot")
```

```{r}
p13 <- ggplot(mb_train_init, aes(x=TEAM_PITCHING_SO, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Pitching Strikeouts", y = "Wins", title="Wins by Strikeouts")
```

```{r}
# Scatterplot with density plot
ggMarginal(p13, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p13, type = "boxplot")
```

```{r}
p14 <- ggplot(mb_train_init, aes(x=TEAM_FIELDING_E, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Fielding Errors", y = "Wins", title="Wins by Errors Commited")
```

```{r}
# Scatterplot with density plot
ggMarginal(p14, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p14, type = "boxplot")
```

```{r}
p15 <- ggplot(mb_train_init, aes(x=TEAM_FIELDING_DP, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm) + 
  labs(x = "Fielding Double Plays", y = "Wins", title="Wins by Defensive Double Plays")
```

```{r}
# Scatterplot with density plot
ggMarginal(p15, type = "density")
```

```{r}
# Scatterplot with boxplot
ggMarginal(p15, type = "boxplot")
```

## Data Exploration Notes

- Very few entries for Batting HBP, consider ignoring that attribute all together
- TEAM_PITCHING_H (consider removing everything above 2000 ... 1620 is 10 hits allowed for EVERY game)
- TEAM_PITCHING_BB (consider removing everything above 1000)
- TEAM_PITCHING_SO (consider removing everything above 2000)
- Apparently there is a team with 0 wins, see how many rows exist like that

```{r}
# Consider all hits as one variable: total bases

# First, identify the number of singles as Hits represents all hit types combined
mb_train_init$TEAM_BATTING_1B <- mb_train_init$TEAM_BATTING_H - mb_train_init$TEAM_BATTING_2B - mb_train_init$TEAM_BATTING_3B - mb_train_init$TEAM_BATTING_HR

mb_train_init$TOTAL_BASES <-  mb_train_init$TEAM_BATTING_1B 

mb_train_init$TOTAL_BASES <- (mb_train_init$TEAM_BATTING_2B * 2) + mb_train_init$TOTAL_BASES

mb_train_init$TOTAL_BASES <- (mb_train_init$TEAM_BATTING_3B * 3) + mb_train_init$TOTAL_BASES

mb_train_init$TOTAL_BASES <- (mb_train_init$TEAM_BATTING_HR * 4) + mb_train_init$TOTAL_BASES

head(mb_train_init)
```

```{r}
ggplot(mb_train_init, aes(x=TOTAL_BASES, y=TARGET_WINS)) + 
  geom_point()+
  geom_smooth(method=lm)
```


```{r}
# How many instances have 0 wins
zero_wins <- subset(mb_train_init, TARGET_WINS == 0)

head(zero_wins)
# Appears to be 1 instance has zero wins
# Looking at the data, appears bogus
```

```{r}
# Remove really bad pitching
bad_pitching <-subset(mb_train_init, TEAM_PITCHING_H > 2000)

summary(bad_pitching)

dim(bad_pitching)
# Result is 257 entries, that seems like a lot of bad pitching
```

```{r}
# Most wins in a season is 116, so let's see how many instance exist above that number
too_many_wins <- subset(mb_train_init, TARGET_WINS > 116)

summary(too_many_wins)

dim(too_many_wins)
# Answer is 17
```

```{r linear_model}
# Note: I have not included BATTING HBP because this causes over 2000 observations to be removed
model <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR +
              TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_init)

summary(model)
```

```{r accuracy_assessment}

sigma(model) / mean(mb_train_init$TARGET_WINS)

```


```{r}
# Note: Only include significant variables
model_sig <- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B +
              TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_init)

summary(model_sig)
```

```{r}

sigma(model_sig) / mean(mb_train_init$TARGET_WINS)

```

```{r}
# Replace all hits as total bases
model_tb <- lm(TARGET_WINS ~ TOTAL_BASES +
              TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_init)

summary(model_tb)
# Did not prove helpful, but leave for now
```

```{r}

sigma(model_tb) / mean(mb_train_init$TARGET_WINS)

```



The lower the RSE, the more accurate the model

Overall, the F-statistic p-value: < 2.2e-16, which is highly significant

Added all the potential predictor variables, so the results show these variables are significant based on checking to see if t-value is significantly different from zero ... those that match significance

TEAM_FIELDING_E  -0.17204    0.04140  -4.155 5.08e-05 ***
TEAM_FIELDING_DP -0.10819    0.03654  -2.961  0.00349 **
TEAM_BATTING_3B  -0.10118    0.07751  -1.305  0.19348 
TEAM_BATTING_BB  -4.45969    3.63624  -1.226  0.22167
TEAM_BASERUN_SB   0.03304    0.02867   1.152  0.25071
TEAM_BATTING_HBP  0.08247    0.04960   1.663  0.09815
TEAM_PITCHING_BB  4.51089    3.63372   1.241  0.21612 


# So let's start removing some data from the data based on questionable data

- Remove the BATTING HBP column
- Remove all entries with too many wins
- Remove all entries with too many pitching hits allowed
- Remove the instance with zero wins

```{r}

# Drop column for Batting HBP
mb_train_init_clean <- subset(mb_train_init, select = -c(TEAM_BATTING_HBP)) 

# Remove entries with too many wins
mb_train_init_clean <- subset(mb_train_init_clean, TARGET_WINS <= 116)

# Remove entries with zero wins (1 total)
mb_train_init_clean <- subset(mb_train_init_clean, TARGET_WINS != 0)

# Remove entries with too many hits allowed
mb_train_init_clean <- subset(mb_train_init_clean, TEAM_PITCHING_H < 2000)

summary(mb_train_init_clean)

dim(mb_train_init_clean)
```


## Split the initial training data into train and validation

```{r}
# https://machinelearningmastery.com/machine-learning-in-r-step-by-step/

# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- createDataPartition(mb_train_init_clean$TARGET_WINS, p=0.80, list=FALSE)
# select 20% of the data for validation
mb_valid_clean <- mb_train_init_clean[-validation_index,]
# use the remaining 80% of data to training and testing the models
mb_train_clean <- mb_train_init_clean[validation_index,]

```

## Attempt linear model on cleaned data

```{r}
model_clean <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR +
              TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean)
```

```{r}
sig <- sigma(model_clean)

sig

sig / mean(mb_train_clean$TARGET_WINS)
```

## Stepwise improvement of cleaned dataset

```{r}
# STEP 1
# Removed TEAM_BATTING_SO
model_clean_st1 <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st1)
```

```{r}
sig <- sigma(model_clean_st1)

sig

sig / mean(mb_train_clean$TARGET_WINS)

```

```{r}
# STEP 2
# Removed TEAM_PITCHING_H
model_clean_st2 <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st2)
```

```{r}
sig <- sigma(model_clean_st2)

sig

sig / mean(mb_train_clean$TARGET_WINS)

```


```{r eval=FALSE}
AIC(model_clean_st2)
```

```{r eval=FALSE}
BIC(model_clean_st2)
```

```{r}
# STEP 3
# Removed TEAM_BATTING_HR
model_clean_st3 <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st3)
```

```{r}
sig <- sigma(model_clean_st3)

sig

sig / mean(mb_train_clean$TARGET_WINS)

```

Continually the best result is Step 3

```{r}
# STEP 4
# Removed TEAM_PITCHING_BB
model_clean_st4 <- lm(TARGET_WINS ~ TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st4)
```

```{r}
sig <- sigma(model_clean_st4)

sig

sig / mean(mb_train_clean$TARGET_WINS)

```

```{r}
plot(model_clean_st4)
```

```{r}
# Graph residuals

ggplot(data=model_clean_st3, aes(model_clean_st3$residuals)) +
  geom_histogram(binwidth = 1, color = "black", fill = "purple4") +
  theme(panel.background = element_rect(fill = "white"),
        axis.line.x=element_line(),
        axis.line.y=element_line()) +
  ggtitle("Histogram for Model Residuals")
```

```{r}
# Run the step wise approach using the stepAIC function
library(MASS)
stepmodel <- stepAIC(model_clean, direction = c("both"), trace = FALSE)

summary(stepmodel)

sig <- sigma(stepmodel)

sig

sig / mean(mb_train_clean$TARGET_WINS)

plot(stepmodel)
```

## Impute missing data
https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/

Using package MICE

```{r}
library(mice)
```

```{r eval=FALSE}

# SKIP THIS SECTION

# Impute the missing data
mb_train_init_imputed <- mice(mb_train_init, m=5, maxit=50, method='pmm', seed=500)


summary(mb_train_init_imputed)

mb_train_imp_2 <- complete(mb_train_init_imputed,2)

# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- createDataPartition(mb_train_imp_2$TARGET_WINS, p=0.80, list=FALSE)
# select 20% of the data for validation
mb_valid_imp <- mb_train_imp_2[-validation_index,]
# use the remaining 80% of data to training and testing the models
mb_train_imp <- mb_train_imp_2[validation_index,]
```


```{r eval=FALSE}
# output one of the imputed dataframes
complete_data_2 <- complete(mb_train_init_imputed,2)

head(complete_data_2)
```

```{r eval=FALSE}
head(mb_train_init)
```

```{r eval=FALSE}
# build predictive model
fit <- with(data=mb_train_init_imputed, exp = lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + 
                                              TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BASERUN_SB + 
                                              TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_BB + 
                                              TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, 
                                            data=mb_train_clean,
                                            na.action = na.omit))

summary(fit)
```

```{r eval=FALSE}
# combine results of all 5 models
pooled <- pool(fit)
summary(pooled)
```

```{r eval=FALSE}
# https://stackoverflow.com/questions/52713733/how-to-use-predict-function-with-my-pooled-results-from-mice


# Copy one of the fitted lm models fit to
#   one of the imputed datasets
pooled_lm = fit$analyses[[1]]
# Replace the fitted coefficients with the pooled
#   estimates (need to check they are replaced in
#   the correct order)
pooled_lm$coefficients = summary(pooled)$estimate

# Predict - predictions seem to match the
#   pooled coefficients rather than the original
#   lm that was copied
# predict(fit$analyses[[1]], newdata = nhanes)
wins_pred_imp <- predict(pooled_lm, newdata = mb_valid_imp)

wins_pred_imp

actual_preds_imp <- data.frame(cbind(actuals=mb_valid_imp$TARGET_WINS, predicteds=wins_pred_imp))

# actual_preds_imp <- subset(actual_preds_imp, predicteds > 0)

actual_preds_imp
```

```{r eval=FALSE}
correlation_accuracy <- cor(actual_preds_imp)

correlation_accuracy
```

```{r eval=FALSE}
mape <- MAPE(actual_preds_imp$predicteds, actual_preds_imp$actuals)

# lower is better
# mape 0.1397487 after removal of negative wins
# 0.3172925 with the negative scores
mape
```

## Run against evaluation data
```{r}
# Read in CSV file of evaluation data
mb_eval <- read.csv("moneyball-evaluation-data.csv")
```

```{r}
# Dimensions of the dataset
dim(mb_eval)
```

```{r}
# list types of each attribute
sapply(mb_eval, class)
```

```{r}
# first five rows
head(mb_eval)
```

```{r}
# summary
summary(mb_eval)
```

```{r}
# Impute the date for the missing validation data
# Impute the missing data
mb_valid_clean_imputed <- mice(mb_valid_clean, m=5, maxit=50, method='pmm', seed=500)


summary(mb_valid_clean_imputed)

mb_valid_clean_imputed_2 <- complete(mb_valid_clean_imputed,2)

# create a list of 80% of the rows in the original dataset we can use for training
#validation_index <- createDataPartition(mb_train_imp_2$TARGET_WINS, p=0.80, list=FALSE)
# select 20% of the data for validation
#mb_valid_imp <- mb_train_imp_2[-validation_index,]
# use the remaining 80% of data to training and testing the models
#mb_train_imp <- mb_train_imp_2[validation_index,]
```


```{r eval=FALSE}
# output one of the imputed dataframes
complete_data_2 <- complete(mb_train_init_imputed,2)

head(complete_data_2)
```

```{r}
library(Metrics)
library(MLmetrics)
# switched mb_eval to mb_valid_clean - I want to see how good predictions are on known target wins

mb_valid_clean <- mb_valid_clean[complete.cases(mb_valid_clean), ]

wins_pred <- predict(model_clean_st3, mb_valid_clean_imputed_2)

# wins_pred

wins_pred_neg <- subset(wins_pred, wins_pred < 0)

wins_pred_neg

rmse <- rmse(mb_valid_clean_imputed_2$TARGET_WINS, wins_pred)

rmse

mape <- MAPE(mb_valid_clean_imputed_2$TARGET_WINS, wins_pred)

mape
```

Need to add code to output the results at runtime: These are previous results

- RMSE: 15.09709

- MAPE: 0.2541715

```{r}
wins_pred <- predict(stepmodel, mb_valid_clean_imputed_2)

# wins_pred

rmse <- rmse(mb_valid_clean_imputed_2$TARGET_WINS, wins_pred)

rmse

mape <- MAPE(mb_valid_clean_imputed_2$TARGET_WINS, wins_pred)

mape
```

Need to add code to output the results at runtime: These are previous results

- RMSE: 15.09709

- MAPE: 0.2541715

```{r}
# http://r-statistics.co/Linear-Regression.html
actual_preds <- data.frame(cbind(actuals=mb_valid_clean_imputed_2$TARGET_WINS, predicteds=wins_pred))
correlation_accuracy <- cor(actual_preds)

correlation_accuracy

# actual_preds

```





## Extra attempts at initial model

```{r eval=FALSE}
# STEP 2 TB (using Total Bases instead of hits individually)
# Removed TEAM_PITCHING_H
model_clean_st2tb <- lm(TARGET_WINS ~ TOTAL_BASES +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st2tb)
```

```{r eval=FALSE}

sigma(model_clean_st2tb) / mean(mb_train_clean$TARGET_WINS)

```

```{r eval=FALSE}
# STEP 3 TB
# Removed TEAM_PITCHING_BB
model_clean_st3tb <- lm(TARGET_WINS ~ TOTAL_BASES +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st3tb)
# Got worse
```

```{r eval=FALSE}

sigma(model_clean_st3tb) / mean(mb_train_clean$TARGET_WINS)

```

```{r eval=FALSE}
# STEP 3 - Got worse
# Removed TEAM_BATTING_HR
model_clean_st3 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B +
              TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BASERUN_CS +
              TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train_clean,
            na.action = na.omit)

summary(model_clean_st3)
```

```{r eval=FALSE}

sigma(model_clean_st3) / mean(mb_train_clean$TARGET_WINS)

```


```{r eval=FALSE}

fitted(model_clean)

```


```{r model_2, eval=FALSE}
model_2 <- lm(TARGET_WINS ~ TEAM_BATTING_3B + TEAM_BATTING_BB + TEAM_BASERUN_SB + TEAM_BATTING_HBP + 
                TEAM_PITCHING_BB + TEAM_FIELDING_E + TEAM_FIELDING_DP, 
              data=mb_train)

summary(model_2)
```

```{r accuracy_assessment_2, eval=FALSE}

sigma(model_2) / mean(mb_train$TARGET_WINS)

```

Got worse ... ha, I knew some of the hitting would need to be included

Add up all the hits for total bases

http://www.philsbaseball.com/Articles/2010_to_2014/2014/September/total_base_percentage.php#:~:text=Here%20is%20the%20formula%3A%20Total,4)%20by%20at%2Dbats.
Here is the formula: Total Bases + walks + hit-by-pitches + stolen bases – caught stealing divided by plate appearances. 

```{r model_TB, eval=FALSE}
model_TB <- lm(TARGET_WINS ~ TOTAL_BASES +
              TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_BATTING_HBP +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train)

summary(model_TB)

sigma(model_TB) / mean(mb_train$TARGET_WINS)
```

```{r eval=FALSE}
mb_train$TOTAL_BASES_PLUS <- mb_train$TEAM_BATTING_BB + mb_train$TOTAL_BASES

head(mb_train)
```

```{r model_TB_plus, eval=FALSE}
model_TB_plus <- lm(TARGET_WINS ~ TOTAL_BASES_PLUS +
              TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_BATTING_HBP +
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + 
              TEAM_FIELDING_DP, 
            data=mb_train)

summary(model_TB_plus)

sigma(model_TB_plus) / mean(mb_train$TARGET_WINS)
```